---
title: "Lab4"
author: "Juehan Wang"
date: "9/17/2021"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(data.table)
library(cowplot)
library(leaflet)
library(ggplot2)
library(Hmisc)
```

```{r data-read, include = FALSE, cache = TRUE}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

## 2. Prepare the data

```{r data-prep, cache=TRUE}
#Remove temperatures less than -17C
met <- met[met$temp >= -17]
#Make sure there are no missing data in the key variables coded as 9999, 999, etc
met[,range(temp,na.rm = TRUE)]
met[,range(rh,na.rm = TRUE)]
met[,range(wind.sp,na.rm = TRUE)]
met[,range(vis.dist,na.rm = TRUE)]
met[,range(dew.point,na.rm = TRUE)]
met[,range(lat,na.rm = TRUE)]
met[,range(lon,na.rm = TRUE)]
met[,range(elev,na.rm = TRUE)]

met[elev == 9999.0, elev :=NA]

#Generate a date variable using the functions as.Date() (hint: You will need the following to create a date paste(year, month, day, sep = "-")).
met[, ymd := as.Date(paste(met$year, met$month, met$day, sep = "-"))]

met[,table(week(ymd))]

#Using the data.table::week function, keep the observations of the first week of the month.
met <- met[week(ymd) ==31]

#Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
met_avg <- met[,.(
  temp      = mean(temp,na.rm=TRUE),
  rh        = mean(rh,na.rm=TRUE),
  wind.sp   = mean(wind.sp,na.rm=TRUE),
  dew.point = mean(dew.point,na.rm=TRUE),
  vis.dist  = mean(vis.dist,na.rm=TRUE),
  lat       = mean(lat),
  lon       = mean(lon), 
  elev      = mean(elev,na.rm=TRUE)
), by=c("USAFID")]

#Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
met_avg[lat >= 39.71 & lon <= -98, region := "Northwest"]
met_avg[lat < 39.71 & lon <= -98, region := "Sorthwest"]
met_avg[lat >= 39.71 & lon > -98, region := "Northeast"]
met_avg[lat < 39.71 & lon > -98, region := "Sortheast"]

met_avg[, region2 := fifelse(lat >= 39.71 & lon <= -98, "Northwest",
        fifelse(lat < 39.71 & lon <= -98, "Sorthwest",
                fifelse(lat >= 39.71 & lon > -98, "Northeast",
                        fifelse(lat < 39.71 & lon > -98, "Sortheast", NA_character_))))]

met_avg[, table(region, useNA = "always")]

#Create a categorical variable for elevation as in the lecture slides
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]
```

## 3. Use geom_violin to examine the wind speed and dew point temperature by region

```{r violin-temp}
#You saw how to use geom_boxplot in class. Try using geom_violin instead (take a look at the help). (hint: You will need to set the x aesthetic to 1)

#Use facets
#Make sure to deal with NA category
#Describe what you observe in the graph
ggplot(met_avg, mapping = aes(y = wind.sp, x = 1)) +
  geom_violin() +
  facet_grid(~region)
```

## 4. Use geom_jitter with stat_smooth to examine the association between dew point temperature and wind speed by region

```{r}
#Colour points by region
#Make sure to deal with NA category
#Fit a linear regression line by region
#Describe what you observe in the graph
met_avg[!is.na(wind.sp) & !is.na(dew.point)] %>%
  ggplot(mapping = aes(x = dew.point, y = wind.sp, color = region, linetype = region)) + 
  geom_jitter() + 
  stat_smooth(formula = y~x, method = lm, se = FALSE, col = "black")
```

## 5. Use geom_bar to create barplots of the weather stations by elevation category coloured by region

```{r}
met_avg[!is.na(elev_cat) & elev_cat != "clear"] %>%
  ggplot() + 
  geom_bar(mapping = aes(x = elev_cat, fill = region), position = "dodge")
```

## 6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars

```{r plot-sumstat}
#Make sure to remove NA
#Use fun.data=“mean_sdl” in stat_summary
ggplot(
  met_avg[!is.na(wind.sp) & !is.na(dew.point)],
  mapping = aes(x = region, y = wind.sp)) +
  stat_summary(fun.data = "mean_sdl") +
  stat_summary(fun.data = "mean_sdl", geom = "errorbar")
```

## 7. Make a map showing the spatial trend in relative h in the US

```{r}
#Add another layer of stats_summary but change the geom to "errorbar" (see the help).
#Describe the graph and what you observe
#Dew point temperature is…
#Wind speed is…
temp.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)

leaflet(met_avg) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(
    lat = ~lat, lng=~lon,
                                                  # HERE IS OUR PAL!
    label = ~paste0(round(rh,2), ' rh'), color = ~ temp.pal(rh),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=temp.pal, values=met_avg$rh,
          title='Temperature, C', opacity=1)
```

