---
title: "Lab4"
author: "Juehan Wang"
date: "9/17/2021"
#output:
#    html_document:
#      toc: yes 
#      toc_float: yes 
#      keep_md: yes
#    github_document:
#      keep_html: true
#      html_preview: false
#always_allow_html: true
output:
  github_document:
    html_preview: false
  html_document: false
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(data.table)
library(cowplot)
library(knitr)
library(leaflet)
library(ggplot2)
library(Hmisc)
library(ggforce)
```

## 1. Read in the data

```{r data-read, include = FALSE, cache = TRUE}
# First download and then read in with data.table:fread()
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

## 2. Prepare the data

```{r data-prep, cache=TRUE}
# Remove temperatures less than -17C
met <- met[met$temp >= -17]
# Make sure there are no missing data in the key variables coded as 9999, 999, etc
met[,range(temp,na.rm = TRUE)]
met[,range(rh,na.rm = TRUE)]
met[,range(wind.sp,na.rm = TRUE)]
met[,range(vis.dist,na.rm = TRUE)]
met[,range(dew.point,na.rm = TRUE)]
met[,range(lat,na.rm = TRUE)]
met[,range(lon,na.rm = TRUE)]
met[,range(elev,na.rm = TRUE)]

met[elev == 9999.0, elev :=NA]

# Generate a date variable using the functions as.Date() (hint: You will need the following to create a date paste(year, month, day, sep = "-")).
met[, ymd := as.Date(paste(met$year, met$month, met$day, sep = "-"))]

met[,table(week(ymd))]

# Using the data.table::week function, keep the observations of the first week of the month.
met <- met[week(ymd) ==31]

# Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
met_avg <- met[,.(
  temp      = mean(temp,na.rm=TRUE),
  rh        = mean(rh,na.rm=TRUE),
  wind.sp   = mean(wind.sp,na.rm=TRUE),
  dew.point = mean(dew.point,na.rm=TRUE),
  vis.dist  = mean(vis.dist,na.rm=TRUE),
  lat       = mean(lat),
  lon       = mean(lon), 
  elev      = mean(elev,na.rm=TRUE)
), by=c("USAFID")]

# Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
met_avg[lat >= 39.71 & lon <= -98, region := "Northwest"]
met_avg[lat < 39.71 & lon <= -98, region := "Southwest"]
met_avg[lat >= 39.71 & lon > -98, region := "Northeast"]
met_avg[lat < 39.71 & lon > -98, region := "Southeast"]

met_avg[, region2 := fifelse(lat >= 39.71 & lon <= -98, "Northwest",
        fifelse(lat < 39.71 & lon <= -98, "Southwest",
                fifelse(lat >= 39.71 & lon > -98, "Northeast",
                        fifelse(lat < 39.71 & lon > -98, "Southeast", NA_character_))))]

met_avg[, table(region, useNA = "always")]

# Create a categorical variable for elevation as in the lecture slides
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]
```

## 3. Use geom_violin to examine the wind speed and dew point temperature by region

```{r violin-temp}
# You saw how to use geom_boxplot in class. Try using geom_violin instead (take a look at the help). (hint: You will need to set the x aesthetic to 1)

# Use facets
# Make sure to deal with NA category
# Describe what you observe in the graph
met_avg[!is.na(wind.sp) & !is.na(temp)] %>%
ggplot(met_avg, mapping = aes(y = wind.sp, x = temp)) +
  geom_violin() +
  facet_grid(~region)
```

From the plot we can see that the dew point temperature in southeast is the highest while that in southwest is the lowest and it also has the widest range. Meanwhile, the southwest has the largest wind speed while the southeast has the smallest.

## 4. Use geom_jitter with stat_smooth to examine the association between dew point temperature and wind speed by region

```{r geom_jitter-stat_smooth}
# Colour points by region
# Make sure to deal with NA category
# Fit a linear regression line by region
# Describe what you observe in the graph
met_avg[!is.na(wind.sp) & !is.na(dew.point)] %>%
  ggplot(mapping = aes(x = dew.point, y = wind.sp, color = region, linetype = region)) + 
  geom_jitter() + 
  stat_smooth(formula = y~x, method = lm, se = FALSE, col = "black")
```

It seems that there is only slightly positive association between dew point temperature and wind speed in Southeast, and there seems no association in the other three regions.

## 5. Use geom_bar to create barplots of the weather stations by elevation category coloured by region

```{r geom_bar}
# Bars by elevation category using position="dodge"
# Change colors from the default. Color by region using scale_fill_brewer see this
# Create nice labels on axes and add a title
# Describe what you observe in the graph
# Make sure to deal with NA
met_avg[!is.na(elev_cat) & elev_cat != "clear"] %>%
  ggplot() + 
  geom_bar(mapping = aes(x = elev_cat, fill = region), position = "dodge")
```

In high elevation category, Northeast has the most weather stations while Northwest has the least. In low elevation category, Southeast has the most weather stations while Northwest has the least.

## 6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars

```{r plot-sumstat}
# Make sure to remove NA
# Use fun.data=“mean_sdl” in stat_summary
# Add another layer of stats_summary but change the geom to "errorbar" (see the help).
# Describe the graph and what you observe
# Dew point temperature is…
# Wind speed is…
ggplot(
  met_avg[!is.na(wind.sp) & !is.na(dew.point)],
  mapping = aes(x = region, y = wind.sp)) +
  stat_summary(fun.data = "mean_sdl") +
  stat_summary(fun.data = "mean_sdl", geom = "errorbar")
```

Wind speed is about 3.2($\pm4$) m/s in Southwest, 2.8($\pm3.2$) m/s in Northwest, 1.7($\pm3.3$) m/s in Northeast and 1.8($\pm3.4$) m/s in Southeast.

## 7. Make a map showing the spatial trend in relative h in the US

```{r spatial-trend}
# Add another layer of stats_summary but change the geom to "errorbar" (see the help).
# Describe the graph and what you observe
# Dew point temperature is…
# Wind speed is…

temp.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)

if (knitr::is_html_output()) {
  leaflet(met_avg) %>%
   addProviderTiles('CartoDB.Positron') %>%
   addCircles(
   lat = ~lat, lng=~lon,
                                                
   label = ~paste0(round(rh,2), ' rh'), color = ~ temp.pal(rh),
   opacity = 1, fillOpacity = 1, radius = 500
     ) %>%
   # And a pretty legend
   addLegend('bottomleft', pal=temp.pal, values=met_avg$rh,
           title='Temperature, C', opacity=1)
} else {
  message("Sorry! No HTML.")
}
```

There is an obvious trend of increasing in rh from the east to the west in the US, and rh is also higher along the coastline. Rh in the east is higher than 60 while it is mostly lower than 60 in the west.

## 8. Use a ggplot extension

```{r ggforce}
# Pick and extension (except cowplot) from here and make a plot of your choice using the met data (or met_avg)
# Might want to try examples that come with the extension first (e.g. ggtech, gganimate, ggforce)
ggplot(met_avg, aes(wind.sp, dew.point, colour = region)) +
  geom_point() +
  facet_zoom(y = region == "Southeast")
```

