---
title: "Lab3"
author: "Juehan Wang"
date: "9/10/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com"))
```

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(leaflet)
library(lubridate)
```

### 1. Read the data

```{r}
# Install it only if it doesn't exist
if (!require(data.table)){      # if the package exists, it will be TRUE
  install.packages("data.table")
}

# Only download the data if it doesn't exist
if (!file.exists("met_all.gz")){
  download.file(
    "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    "met_all.gz",
    method="libcurl",
    timeout = 60
    )
}

met <- data.table::fread("met_all.gz")
```

### 2. Check the dimensions, headers, footers.

```{r}
dim(met)
head(met)
tail(met)
```

### 3. Take a look at the variables.

```{r}
str(met)
```

### 4. Take a closer look at the key variables.

```{r}
table(met$year)
table(met$day)
table(met$hour)
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
#met[met$elev==9999.0] <- NA
  # same as
met[elev == 9999.0, elev :=NA]    # ":=" modify elev with NA
summary(met$elev)
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

The highest weather station is at `r max(met$elev, na.rm = TRUE)` `r ifelse(runif(1) > .5, "ft", "mts")`.

### 5. Check the data against an external data source.

```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

### 6. Calculate summary statistics

```{r}
met[elev==max(elev,na.rm = TRUE)][, summary(wind.sp)]
met[elev==max(elev,na.rm = TRUE)][, summary(temp)]
```

```{r}
met[elev == max(elev, na.rm = TRUE), .(
  temp_wind = cor(temp, wind.sp, use="complete"),
  temp_day = cor(temp, day, use="complete"),
  temp_hour = cor(temp, hour, use="complete"),
  wind_day = cor(wind.sp, day, use="complete"),
  wind_hour = cor(wind.sp, hour, use="complete")
)]
```

### 7. Exploratory graphs

```{r}
hist(met$elev, breaks=100)
hist(met$temp)
hist(met$wind.sp)
```

Look at where the weather station with highest elevation is located.

```{r}
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

Look at the time series of temperature and wind speed at this location. 

```{r}
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)

elev <- elev[order(date)]
head(elev)
```