---
title: "Lab 5"
author: "Juehan Wang"
date: "9/24/2021"
output:
    html_document:
      toc: yes 
      toc_float: yes 
      keep_md: yes
    github_document:
      keep_html: true
      html_preview: false
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(data.table)
library(tidyverse)
library(tidytext)
library(tibble)
library(dplyr)
```

First, download the data.

```{r get-data}
fn <- "mtsamples.csv"
if (!file.exists(fn))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv", destfile = fn)

mtsamples<-read.csv(fn)
mtsamples<-as_tibble(mtsamples)
```

### Question 1: What specialties do we have?

We can use count() from dplyr to figure out how many different categories do we have? Are these categories related? overlapping? evenly distributed?

```{r dist-of-specialists}
specialties <- mtsamples %>%
  count(medical_specialty, sort = TRUE)
```

There are `r nrow(specialties)` specialties. Let's take a look at the distributions.

```{r dist1}
ggplot(mtsamples, aes(x = medical_specialty)) +
  geom_histogram(stat = "count") +
  coord_flip()

specialties %>%
  arrange(desc(n)) %>%
  top_n(n, 15) %>%
  knitr::kable()
```

```{r dist2}
ggplot(specialties, aes(x = n, y = fct_reorder(medical_specialty,n))) +
  geom_col()
```

These are not evenly (uniformly) distributed.

### Question 2

Tokenize the the words in the transcription column

Count the number of times each token appears

Visualize the top 20 most frequent words

Explain what we see from this result. Does it makes sense? What insights (if any) do we get?

```{r token-trans, cache=TRUE}
mtsamples %>%
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  top_n(20) %>%
  ggplot(aes(x = n, y = fct_reorder(word,n))) +
    geom_col()
```

The word "patient" seems to be important (duh!), but we observe a lot of stopwords.

### Question 3



What do we see know that we have removed stop words? Does it give us a better idea of what the text is about?

```{r token-trans-wo-stop, cache=TRUE}
# Redo visualization but remove stopwords before
mtsamples %>%
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = "word") %>%
  top_n(20) %>%
  ggplot(aes(x = n, y = fct_reorder(word,n))) +
    geom_col()
```

```{r}
# Bonus points if you remove numbers as well
```

```{r}

```






